---
title: "Modelling acoustics in the digital thigh"
author: "R. Grdina"
date: "08/23/2015"
output: html_document
---

Here we copy W. R. Bauer's back-of-the-envelope acoustic transmission model developed for the BrainWeb phantom to his digital thigh phantom. As he did for a voxel, we'll use ray traversal through pixels. A "ray" is represented by the sequence of pixels a straight line between a specified transmitter and receiver intersects. "As ultrasound traverses the ray, it accumulates losses due to absorption in voxel interiors and reflection at voxel boundaries." We'll ignore reflections and refractions for the time being and  instead consider only the inverse square loss suffered between the transmitter and receiver. In decibels this loss would be $10\:log_{10}(r^2)$ where $r$ is distance between the endpoints.

Acoustic properties for the phantom's 8 tissue types were estimated [here] in csv form [here](../data/acoustic_properties_thigh.csv). In addition to the tissue identifier, the table includes the density, velocity, and acoustic impedance of each tissue. This last is the product of the first two, i.e., $Z = \rho \cdot c$, where $Z$ represents impedance, $\rho$ density, and $c$ velocity, respectively.

```{r}
properties <- read.table("../data/acoustic_properties_thigh.csv", sep=",", header=TRUE)
rho <- properties$rho
speed <- properties$speed
alpha <- runif(8, properties[,"alpha.low"], properties[,"alpha.high"])
```

Absorption coefficients, alpha, are given in units of $dB/m/MHz$. Since the phantom's voxels are 1 mm on a side, we'll convert to $dB/mm/MHz$. 

```{r}
alpha <- alpha/1000
```

The other necessary quantities are transmission coefficients, also expressed in decibels of loss. These are most easily arranged in a matrix. The general formula for transmitted to (normally incident) power is$$10\:log_{10}\left(\frac{4Z_1Z_2}{(Z_1 + Z_2)^2}\right),$$where $Z_i = \rho_i c_i.$ There is no reflection between tissues with identical properties, hence some matrix elements will be zero.

```{r}
Z <- rho*speed
rc <- matrix(0, 8, 8)
for(i in 1:8){
  for(j in 1:8){
    if(Z[i] != Z[j]){
      rc[i,j] <- 10*log10(4*Z[i]*Z[j]/(Z[i]+Z[j])^2)
    }
  }
}
```

We'll use a small section of a sagittal slice for testing. A voxelized ray is shown in white. (The code which produced the figure is hidden for simplicity.)

```{r echo=FALSE, fig.align='center', fig.show='hold'}
source("../R/utilities.R")
fname <- "../data/subject04_crisp_v.rawb"
phantom <- readBin(fname, what="raw", n=362*434*362, size=1, signed=FALSE, endian="big")
dim(phantom) <- c(362, 434, 362)
slice <- phantom[100,,]
slice <- as.integer(slice)
dim(slice) <- c(434, 362)
rm(phantom)
section <- slice[280:350, 250:310]
# section[10,11] # gray
# section[30, 39] # muscle/skin next to scalp
temp <- findVoxelCrossings(c(0,10-.5,11-.5), c(0,30-.5,39-.5))
ray <- matrix(c(0,10,11), 1, 3)
for(n in seq(1, nrow(temp))){
  nxt <- ray[nrow(ray),]
  k <- temp[n,4]
  nxt[k] <- 1 + nxt[k]
  ray <- rbind(ray, nxt)
}
disp_slice(section, "Voxelized ray from gray matter to scalp")
points(ray[,2], ray[,3], pch=15, col="white")
```

The variable, `section`, below codes the sagittal section shown in the figure. The voxel sequence is contained the variable, `ray`. Since the pictured section is in the y, z plane, the first column of `ray` is irrelevant. We first obtain the sequence of tissue types along the ray.

```{r}
tissues <- sapply(1:nrow(ray), function(n){section[ray[n,2], ray[n,3]]})
```

We use the tissue sequence to calculate absorption loss.

```{r}
abs_loss <- sum(alpha[tissues])
```

Tissue transitions are used to calculate reflection losses.

```{r}
refl_loss <- -sum(sapply(2:length(tissues), function(n)rc[tissues[n-1]+1, tissues[n]+1]))
```

Finally, r-squared loss.

```{r}
rsq_loss <- 10*log10((30-10)^2 + (39-11)^2)
c(absorption=abs_loss, reflection=refl_loss, r_squared=rsq_loss)
```

R-squared loss is clearly dominant. To overcome the ~30 dB loss per square mm, a collection area of about 1000, or ~32x32, square mm would be required.

We'll consolidate the above code in appropriate utilities and make more thorough and representative runs in the near future. The code used here is visible in the associated [Rmd file](acoustic_1.Rmd).