---
title: "Distribution of lines through a pixel"
output: html_document
---

```{r echo=FALSE}
source("../R/utilities.R")
```

Assume that$$Sb=ToF+\eta$$where $b$ is a $2^n \times 2^n$ array representing slowness values in a corresponding array of pixels, $S$ is a $2^{4n} \times 2^{4n}$ matrix whose elements represent the length of flight of a given transmitter-to-receiver path (row) through a given pixel (column,) and $\eta$ represents time of flight measurement errors represented as i.i.d. normally distributed random variables with zero mean.

In an MCMC or a hillclimbing algorithm, $b$ would be varied incrementally and error recomputed. One possibility is to vary a single element of $b$ at a time, (e.g., Gibbs sampling in MCMC.) If $\Delta b$ represents the change in $b$ and $\eta'$ the new error,$$S(b + \Delta b) = ToF + \eta', \\ Sb + S\Delta b = ToF + \eta', \\ \eta' = S\Delta b + (Sb - ToF), \\ \eta' = S\Delta b + \eta.$$

Thus, $S\Delta b$ represents the change in putative error. If only a single element of $b$ were changed per step of an algorithm, the computation of $S\Delta b$ would involve only one column of $S$ and only those rows of $S$ corresponding to paths which intersect the pixel corresponding to the changed element of $b$. It is therefore of some interest to know how many paths intersect an average pixel and how that number increases with $n.$

We estimate the distribution--mean, standard deviation, and histogram--of such numbers for $n=5, 6, 7$ (32x32, 64x64, 128x128, pixels respectively,) by choosing 1000 pixels at random and calculating the number of paths which intersect each one.

Roughly, the mean is $1.25*2^n$ and the standard deviation is $45\%$ of the mean.

```{r echo=FALSE}
mcrun <- function(n, trials){
  # The pixel array will be 2^n by 2^n, with z (horizontal) and x (vertical)
  # coordinates ranging between 0 and 2^n.
  # Transmitters will be located at z=0, x = .5 + 0:(2^n-1)
  # and receivers at z=2^n, x = .5 + 0:(2^n-1)
  zxmit <- 0
  zrcv <- 2^n
  convenience <- 0:(2^n - 1)
  xmit <- xrcv <- 0.5 + convenience
  single_trial <- function(i){
    # Choose a pixel at random
    cornerLL <- sample(convenience, 2) # random lower left corner
    cornerUR <- cornerLL + 1 # add 1 to both coordinates for upper right
    # Find the paths which cross the pixel and return their number
    temp <- findLinesCrossingPixel(cornerLL, cornerUR, xmit, zxmit, xrcv, zrcv)
    sum(sapply(temp, function(x)length(x))) # sum the lines originating from each transmitter 
  }
  sapply(1:trials, single_trial)
}
```

```{r echo=FALSE, fig.align='center', fig.show='hold'}
nis5 <- mcrun(5, 1000)
hist(nis5, xlab="Lines per pixel", main="Lines per Pixel, n=5\n(32x32 pixels)", breaks=20, probability=TRUE, col="lightblue")
lines(density(nis5), lwd=3, col="red", lty=2)
legend('topleft', c(paste0("mean: ", round(mean(nis5))), paste0("sd:      ", round(sd(nis5)))))

nis6 <- mcrun(6, 1000)
hist(nis6, xlab="Lines per pixel", main="Lines per Pixel, n=6\n(64x64 pixels)", breaks=20, probability=TRUE, col="lightblue")
lines(density(nis6), lwd=3, col="red", lty=2)
legend('topleft', c(paste0("mean: ", round(mean(nis6))), paste0("sd:      ", round(sd(nis6)))))

nis7 <- mcrun(7, 1000)
hist(nis7, xlab="Lines per pixel", main="Lines per Pixel, n=7\n(128x128 pixels)", breaks=20, probability=TRUE, col="lightblue")
lines(density(nis7), lwd=3, col="red", lty=2)
legend('topleft', c(paste0("mean: ", round(mean(nis7))), paste0("sd:      ", round(sd(nis7)))))
```