---
title: "Revised test suite from thigh phantom"
author: "W. Bauer"
date: "October 24, 2015"
output: html_document
---

This note adds to the suite of test data set constructed [here](test_suite_thigh.html). The earlier set assumed a probe width of 3 cm which is incorrect. The correct value of 6 cm is used here. 

As before, the first task was to identify rectangular sections. This was a trial and error process, performed off-line. Resulting sections are shown below in various colors. All are 30 mm in height, which is approximately that of a 128 element ultrasound probe.

```{r fig.align='center', message=FALSE, echo=FALSE, fig.width=8}
source("../R/inkscape_utils.R")

phantom <- importIPhantom("../data/thigh_inkscape.csv")

par(mfrow=c(1,2))
plotIPhantom(phantom, main="Scans 1-4")
scan_1 <- computeRectangle(phantom, c(100,82), pi/4, 127*.46)
overlayRectangle(scan_1, lwd=3)
scan_2 <- computeRectangle(phantom, c(108,82), pi/4 + pi/8, 127*.46)
overlayRectangle(scan_2, lwd=3, border=2)
scan_3 <- computeRectangle(phantom, c(90, 77), 0, 127*.46)
overlayRectangle(scan_3, lwd=3, border=3)
scan_4 <- computeRectangle(phantom, c(180, 110), pi/8, 127*.46)
overlayRectangle(scan_4, lwd=3, border=4)

plotIPhantom(phantom, main="Scans 6-10")
scan_5 <- computeRectangle(phantom, c(135,150), pi/2+pi/4+pi/16, 127*.46)
overlayRectangle(scan_5, lwd=3, border=1)
scan_6 <- computeRectangle(phantom, c(130,125), pi/2+pi/4-pi/8, 127*.46)
overlayRectangle(scan_6, lwd=3, border=2)
scan_7 <- computeRectangle(phantom, c(121,125), pi/2-pi/16, 127*.46)
overlayRectangle(scan_7, lwd=3, border=3)
scan_8 <- computeRectangle(phantom, c(150,135), pi/4+pi/16, 127*.46)
overlayRectangle(scan_8, lwd=3, border=4)
par(mfrow=c(1,1))
```

Controlling an image's height in pixels is a bit tricky. The following checks a calculation to do so.

```{r echo=FALSE}
local({
# Utility to compute base and height in pixels, given a rectangle and gridsize.
# It was used to ensure that heights in pixels would be powers of 2 when images
# were pixelated. The height in pixels will be equal to the number of transducers
# when a system matrix is calculated via genS or its C counterpart.
pixi <- function(r, gridsize){
  # assuming rectangle corners are arranged in clockwise order,
  # compute oriented base and height
  bv <- r[2,]-r[1,]
  hv <- r[4,]-r[1,]
  # rectangle dimensions in integer multiples of gridsize
  b <- ceiling(.5+sqrt(sum(bv^2))/gridsize)
  h <- ceiling(.5+sqrt(sum(hv^2))/gridsize)
  # gridsize increments for base and height
  bi <- gridsize*bv/(sqrt(sum(bv^2)))
  hi <- gridsize*hv/(sqrt(sum(hv^2)))
  c(h, b)
}

# Utility to calculate gridsize in mm which will give a certain height, h, in pixels
# for a given rectangl, r.
calcGridSize <- function(r, h){
  hv <- r[4,] - r[1,]
  lhv <- sqrt(sum(hv^2))
  mean(c(lhv/(h-.5), lhv/(h-1.5)))
}

scans <- paste0("scan_",1:8)
target_height <- integer()
actual_height <- integer()
grid_size <- numeric()
for(n in 1:8){
  scan <- get(scans[n], envir=globalenv())
  for(res in 2^(7:4)){
    target_height <- c(target_height, res)
    g <- calcGridSize(scan, res)
    grid_size <- c(grid_size, g)
    actual_height <- c(actual_height, pixi(scan, g)[1])
  }
}
data.frame(scan=rep(scans,each=4), target_height=target_height, actual_height=actual_height, grid_size_mm = grid_size)
})
```

Several minutes of run time were required to produce pixelated images. Hence, the suite was produced off line by the following code. Pixels were labeled by tissue type so that speeds of sound can be varied dynamically. The images themselves are in this project's `data` subdirectory and are identified by the prefix, "wide", the scan number, and the resolution. For instance, `wide_scan_6_64.csv` corresponds to the 6^th^ rectangular section with a grid size corresponding to 64 transducers spread over a 30 cm height. Scan 6 is shown at 4 resolutions below.

```{r eval=FALSE}
datadir = "../data"
scans <- paste0("scan_",1:8)
for(n in 1:8){
  scan <- get(scans[n])
  for(res in 2^(4:7)){
    gridsize <- calcGridSize(scan, res)
    img <- pixelate(phantom, scan, gridsize)
    colnames(img) <- as.character(signif(gridsize*(1:ncol(img)-1),4))
    outf <- paste0(datadir,"/wide_",scans[n],"_",res,".csv")
    write.csv(as.data.frame(img),outf, row.names=FALSE)
    writeLines(outf)
  }
}
```


```{r echo=FALSE, fig.align='center', fig.show='hold', fig.height=4}

readWideScan <- function(scan, resolution, datadir="../data"){
  temp <- as.matrix(read.csv(paste0(datadir,"/wide_scan_",scan,"_",resolution,".csv"), 
                             as.is=TRUE, comment.char = "#"))
  colnames(temp) <- gsub("X","",colnames(temp))
  temp
}
for(res in 2^(7:4)){
  img <- readWideScan(6,res,datadir="../data")
  image(x = .46 * seq(0, nrow(img)-1)*128.0/res,
        y = .46 * seq(0, ncol(img)-1)*128.0/res,
        z = img, xlab="x (mm)", ylab="y (mm)", 
        main = paste0("Wide scan 6, ", res, " transducers"), asp=1)
}

```

