---
title: "Summary 7/2/2015"
author: "gina and bill"
date: "07/02/2015"
output: html_document
---

### Transducers per pixel to minimize null space

We know that if there are n columns of pixels, then the null space of the system matrix has dimension *at least* n-1. It may be more, depending on the number of transducers. It seems that there is always some minimum number of transducers which reduces the null space to its smallest size. However, the relationship is not simple.

```{r echo=FALSE, fig.align='center',fig.show='hold'}
temp <- matrix( c(
6,4,3,
9,7,6,
10,8,7,
11,9,8,
12,10,9,
15,11,10,
20,12,11,
18,13,12,
23,14,13),3,9)
plot(temp[2,], temp[1,], type='p', pch=19, xlab="Number of Pixel Columns", ylab="Minimum Number of Transducers", main="Conjectured Minimum Number of Transducers\nto Minimize Null Space")
abline(2,1,col="blue")
text(12,12,"y = x + 2", col="blue")
```

### Two scans are better than one

```{r echo=FALSE, fig.align='center',fig.show='hold',fig.width=9}
source("../R/handy_scanner.R")
par(mfrow=c(1,2))
setup <- handySetup(16, 15, 12)
showSetup(setup,legends=FALSE)
tof <- scanSetup(setup)
myS <- createSij(setup)
setupY <- alignY(setup)
showSetup(setupY,legends=FALSE)
tofY <- scanSetup(setupY)
mySY <- createSij(setupY)
myS <- rbind(myS,mySY)
tof <- c(as.vector(tof),as.vector(tofY))
myb <- solve(t(myS) %*% myS,t(myS) %*% tof)
myb <- matrix(myb,12,12)
par(mfrow=c(1,1))
image(myb,asp=1,col=gray.colors(50))
```

