---
title: "Suite of tests from thigh phantom"
author: "W. Bauer"
date: "October 24, 2015"
output: html_document
---

This note documents construction of a test suite of pixelated images based on a [phantom thigh section](thigh_working_prototype.html). The suite does not involve misalignment or noise. It involves rectangular sections at a total of 8 positions and orientations, each digitized at resolutions (pixel sizes) corresponding to transducer arrays of 16, 32, 64, and 128 elements.

The first task was to identify rectangular sections. This was a trial and error process, performed off-line. Resulting sections are shown below in various colors. All are 30 mm in height, which is approximately that of a 128 element ultrasound probe.

```{r fig.align='center', message=FALSE, echo=FALSE, fig.width=8}
source("../R/inkscape_utils.R")

phantom <- importIPhantom("../data/thigh_inkscape.csv")

par(mfrow=c(1,2))
plotIPhantom(phantom, main="Scans 1-4")
scan_1 <- computeRectangle(phantom, c(100,100), pi/4, 127*.23)
overlayRectangle(scan_1, lwd=3)
scan_2 <- computeRectangle(phantom, c(100,100), pi/4 + pi/8, 127*.23)
overlayRectangle(scan_2, lwd=3, border=2)
scan_3 <- computeRectangle(phantom, c(90, 80), 0, 127*.23)
overlayRectangle(scan_3, lwd=3, border=3)
scan_4 <- computeRectangle(phantom, c(180, 125), pi/8, 127*.23)
overlayRectangle(scan_4, lwd=3, border=4)

plotIPhantom(phantom, main="Scans 6-10")
scan_5 <- computeRectangle(phantom, c(125,140), pi/2+pi/4+pi/16, 127*.23)
overlayRectangle(scan_5, lwd=3, border=1)
scan_6 <- computeRectangle(phantom, c(130,125), pi/2+pi/4-pi/8, 127*.23)
overlayRectangle(scan_6, lwd=3, border=2)
scan_7 <- computeRectangle(phantom, c(130,125), pi/2-pi/16, 127*.23)
overlayRectangle(scan_7, lwd=3, border=3)
scan_8 <- computeRectangle(phantom, c(180,125), pi/4+pi/16, 127*.23)
overlayRectangle(scan_8, lwd=3, border=4)
par(mfrow=c(1,1))
```

Controlling an image's height in pixels is a bit tricky. The following checks a calculation to do so.

```{r echo=FALSE}
local({
# Utility to compute base and height in pixels, given a rectangle and gridsize.
# It was used to ensure that heights in pixels would be powers of 2 when images
# were pixelated. The height in pixels will be equal to the number of transducers
# when a system matrix is calculated via genS or its C counterpart.
pixi <- function(r, gridsize){
  # assuming rectangle corners are arranged in clockwise order,
  # compute oriented base and height
  bv <- r[2,]-r[1,]
  hv <- r[4,]-r[1,]
  # rectangle dimensions in integer multiples of gridsize
  b <- ceiling(.5+sqrt(sum(bv^2))/gridsize)
  h <- ceiling(.5+sqrt(sum(hv^2))/gridsize)
  # gridsize increments for base and height
  bi <- gridsize*bv/(sqrt(sum(bv^2)))
  hi <- gridsize*hv/(sqrt(sum(hv^2)))
  c(h, b)
}

# Utility to calculate gridsize in mm which will give a certain height, h, in pixels
# for a given rectangl, r.
calcGridSize <- function(r, h){
  hv <- r[4,] - r[1,]
  lhv <- sqrt(sum(hv^2))
  mean(c(lhv/(h-.5), lhv/(h-1.5)))
}

scans <- paste0("scan_",1:8)
target_height <- integer()
actual_height <- integer()
grid_size <- numeric()
for(n in 1:8){
  scan <- get(scans[n], envir=globalenv())
  for(res in 2^(7:4)){
    target_height <- c(target_height, res)
    g <- calcGridSize(scan, res)
    grid_size <- c(grid_size, g)
    actual_height <- c(actual_height, pixi(scan, g)[1])
  }
}
data.frame(scan=rep(scans,each=4), target_height=target_height, actual_height=actual_height, grid_size_mm = grid_size)
})
```

Several minutes of run time were required to produce pixelated images. Hence, the suite was produced off line by a function in file `thigh_script.R`. Pixels were labeled by tissue type so that speeds of sound can be varied dynamically. The images themselves are in this project's `data` subdirectory and are identified by scan number and resolution. For instance, `scan_6_64.csv` corresponds to the 6^th^ rectangular section with a grid size corresponding to 64 transducers spread over a 30 cm height. Scan 6 is shown at 4 resolutions below. 

```{r echo=FALSE, fig.align='center', fig.show='hold', fig.height=3}
source("../R/thigh_script.R")
for(res in 2^(7:4)){
  img <- readScan(6,res,datadir="../data")
  image(x = .23 * seq(0, nrow(img)-1)*128.0/res,
        y = .23 * seq(0, ncol(img)-1)*128.0/res,
        z = img, xlab="x (mm)", ylab="y (mm)", 
        main = paste0("Scan 6, ", res, " transducers"), asp=1)
}

```

File `thigh_script.R` contains two utilities whose use is illustrated below.

```{r fig.align='center', fig.show='hold', fig.height=3}
# Read scan 3, resolution 32
img <- readScan(3, 32, datadir="../data")
# Convert tissue identifiers to slowness values
img <- tissue2Slowness(img, datadir = "../data")
# display the result
image(x = 4*.23 * seq(0, nrow(img)-1),
        y = 4* .23 * seq(0, ncol(img)-1),
        z = img, xlab="x (mm)", ylab="y (mm)", 
        main = paste0("Scan 3, ", 32, " transducers (slowness)"), asp=1)
```

