---
title: "Suite of tests from thigh phantom"
author: "W. Bauer"
date: "October 24, 2015"
output: html_document
---

This note is to document construction of a test suite of images, system matrices, and time of flight data based on a [phantom thigh section](thigh_working_prototype.html). The suite will not involve misalignment or noise, at least initially. It will involve rectangular sections at various positions and orientations and digitized at various resolutions (pixel sizes.)

The first task is to identify rectangular sections. This is a trial and error process, and was performed off-line. Resulting sections are shown below in various colors. All are 30 mm in height, which is approximately that of a 128 element ultrasound probe.

```{r fig.align='center', fig.width=4.4, message=FALSE, echo=FALSE}
source("../R/inkscape_utils.R")

phantom <- importIPhantom("../data/thigh_inkscape.csv")


plotIPhantom(phantom, main="Scans 1-4")
scan_1 <- computeRectangle(phantom, c(100,100), pi/4, 127*.23)
overlayRectangle(scan_1, lwd=3)
scan_2 <- computeRectangle(phantom, c(100,100), pi/4 + pi/8, 127*.23)
overlayRectangle(scan_2, lwd=3, border=2)
scan_3 <- computeRectangle(phantom, c(90, 80), 0, 127*.23)
overlayRectangle(scan_3, lwd=3, border=3)
scan_4 <- computeRectangle(phantom, c(180, 125), pi/8, 127*.23)
overlayRectangle(scan_4, lwd=3, border=4)

plotIPhantom(phantom, main="Scans 6-10")
scan_5 <- computeRectangle(phantom, c(125,140), pi/2+pi/4+pi/16, 127*.23)
overlayRectangle(scan_5, lwd=3, border=1)
scan_6 <- computeRectangle(phantom, c(130,125), pi/2+pi/4-pi/8, 127*.23)
overlayRectangle(scan_6, lwd=3, border=2)
scan_7 <- computeRectangle(phantom, c(130,125), pi/2-pi/16, 127*.23)
overlayRectangle(scan_7, lwd=3, border=3)
scan_8 <- computeRectangle(phantom, c(180,125), pi/4+pi/16, 127*.23)
overlayRectangle(scan_8, lwd=3, border=4)

```

The plan is to produce pixelated images of each rectangular section at various resolutions (pixel sizes.) Heights in pixels correspond to the number of transducers in a probe. These should be powers of 2 ranging from 16 to 128. Controlling the height in pixels is a bit tricky. The following checks a calculation to do so.

```{r echo=FALSE}
local({
# Utility to compute base and height in pixels, given a rectangle and gridsize.
# It was used to ensure that heights in pixels would be powers of 2 when images
# were pixelated. The height in pixels will be equal to the number of transducers
# when a system matrix is calculated via genS or its C counterpart.
pixi <- function(r, gridsize){
  # assuming rectangle corners are arranged in clockwise order,
  # compute oriented base and height
  bv <- r[2,]-r[1,]
  hv <- r[4,]-r[1,]
  # rectangle dimensions in integer multiples of gridsize
  b <- ceiling(.5+sqrt(sum(bv^2))/gridsize)
  h <- ceiling(.5+sqrt(sum(hv^2))/gridsize)
  # gridsize increments for base and height
  bi <- gridsize*bv/(sqrt(sum(bv^2)))
  hi <- gridsize*hv/(sqrt(sum(hv^2)))
  c(h, b)
}

# Utility to calculate gridsize in mm which will give a certain height, h, in pixels
# for a given rectangl, r.
calcGridSize <- function(r, h){
  hv <- r[4,] - r[1,]
  lhv <- sqrt(sum(hv^2))
  mean(c(lhv/(h-.5), lhv/(h-1.5)))
}

scans <- paste0("scan_",1:8)
target_height <- integer()
actual_height <- integer()
grid_size <- numeric()
for(n in 1:8){
  scan <- get(scans[n], envir=globalenv())
  for(res in 2^(7:4)){
    target_height <- c(target_height, res)
    g <- calcGridSize(scan, res)
    grid_size <- c(grid_size, g)
    actual_height <- c(actual_height, pixi(scan, g)[1])
  }
}
data.frame(scan=rep(scans,each=4), target_height=target_height, actual_height=actual_height, grid_size_mm = grid_size)
})
```

Several minutes of run time are required to produce pixelated images. Hence, the suite was produced off line by a function in file `thigh_script.R`.

```{r echo=FALSE}
readScan <- function(scan, resolution){
  temp <- as.matrix(read.csv(paste0("../data/scan_",scan,"_",resolution,".csv"), 
                             as.is=TRUE, comment.char = "#"))
  colnames(temp) <- gsub("X","",colnames(temp))
  temp
}


```

