---
title: "Tissue type transitions, thigh section"
author: "W. Bauer"
date: "September 1, 2015"
output: html_document
---

For convenience, I've made a csv file of the pixelated section of "digital thigh" used for illustration [here](thigh_working_prototype.html). This note estimates statistical properties of possible interest in dealing with the problem of incomplete information. In particular, it estimates probabilites that various tissue types appear in adjacent pixels.


```{r fig.align='center', fig.height=3}
img <- as.matrix(read.csv("../data/pixelized_thigh_section.csv", comment.char = "#"))
image(img, xlab="x (180 mm)", ylab="y (30 mm)", asp=dim(img)[2]/dim(img)[1], xaxt='n', yaxt='n', col=gray.colors(length(unique(as.vector(img)))))
```

I'll estimate horizontal (x direction) and vertical adjacency statistics separately. First, vertically, i.e., across rows by column.

```{r}
n <- dim(img)[2]
# All pairs img[i,j], img[i, j+1]
pairs <- cbind(as.vector(img[,-n]), as.vector(img[,-1]))
# Unique pairs
upairs <- unique(pairs)
# Counts
counts <- numeric(nrow(upairs))
for(i in 1:length(counts)){
  counts[i] <- sum(pairs[,1]==upairs[i,1] & pairs[,2]==upairs[i,2])
}
ycounts <- cbind(upairs, counts)
colnames(ycounts) <- c("y", "y+1", "ycount")
ycounts <- ycounts[order(ycounts[,3], decreasing = TRUE), ]
ycounts
```

Horizontally 

```{r}
n <- dim(img)[1]
# All pairs img[i+1,j], img[i, j]
pairs <- cbind(as.vector(img[-n,]), as.vector(img[-1,]))
# Unique pairs
upairs <- unique(pairs)
# Counts
counts <- numeric(nrow(upairs))
for(i in 1:length(counts)){
  counts[i] <- sum(pairs[,1]==upairs[i,1] & pairs[,2]==upairs[i,2])
}
xcounts <- cbind(upairs, counts)
colnames(xcounts) <- c("x", "x+1", "xcount")
xcounts <- xcounts[order(xcounts[,3], decreasing = TRUE), ]
xcounts
```

I've saved the counts as `thigh_y_digraphs.csv` and `thigh_x_digraphs.csv` in the data subdirectory.