---
title: "Using the box phantom scanner"
author: "W. Bauer"
date: "05/12/2015"
output: html_document
---

This is a test and illustration of code written to automate and generalize what was done manually in [Scanning a simulated phantom](sim_scan_1.html). It is also a preliminary look at the effect of misalignment on time-of-flight data.

```{r}
source("../R/box_scanner.R")
```

I'll first create a box phantom and two probes. To keep run times short I'll give the probes only 32 elements.

```{r}
phantom <- newBoxPhantom(dimensions=c(30, 30, 30), 
                         ctr_alcohol=c(8, 15-1.5, 18), r_alcohol=6, 
                         ctr_water=c(24, 16+1.5, 12), r_water=5,
                         alignment_in_world = diag(1,4,4))
xmitr <- newProbe(n=32, spacing = 30/32)
rcvr <- newProbe(n=32, spacing = 30/32)
```

The `alignProbes` function automates much of the manual process. It will align the probes contralaterally on the phantom boundary, at centers determined by a point and a direction. I'll first align them as in  [Scanning a simulated phantom](sim_scan_1.html) using a point, `c(15, 15, 15)` at the phantom center, and the x direction, `c(1, 0, 0)`. Function `plotSectionAndArrays` will display the setup.

```{r fig.align='center'}
temp <- alignProbes(phantom, c(15,15,15), c(1,0,0), xmitr, rcvr)
plotSectionAndArrays(phantom, temp$transmitters, temp$receivers, temp$z, by=2)
```

The value returned from `alignProbes` is a list containing the locations of transmitters and receivers relative to the phantom's internal reference frame, and the world coordinate, z, of the relevant axial section. The function, `doScan`, will simulate a scan, and function `plotScan` will plot the result.

```{r fig.align='center'}
scan <- doScan(phantom, temp$transmitters, temp$receivers)
plotScan(scan)
```

I'll change the alignment of the receiving probe by a slight rotation about the x axis. Function `recalcForScan` will recalculate the information I need for a second scan.

```{r}
rcvr$adjustBy(affineTransform(c(1,0,0), pi/20, c(0,0,0)))
temp <- recalcForScan(phantom, xmitr, rcvr)
scan2 <- doScan(phantom, temp$transmitters, temp$receivers)
plotScan(scan2)
```

The change is clear, though not necessarily informative.