---
title: "Writing system matrix utilities"
author: "W. Bauer"
date: "July 19, 2015"
output: html_document
---

This is a design document for utilities to facilitate study of the system matrix, which is known to have a non-zero kernel, even in favorable configurations.

The basic setup will be a pixelated test image consisting of a matrix of slowness (reciprocal speed of sound) values. R allows objects such as matrices to have attributes. Function `testImage` creates a matrix with attributes, `spacing`, for pixel spacing (dimension of a side in mm,) and `slowness` for the default value of slowness in a pixel. If slowness is not specified, `1/1500` mm/ms is used.

```{r}
source("../R/system_matrix_utils.R")
img <- testImage(32, 32, 1)
```

Representative speeds of sound in tissues of interest are given in the vector, `speed`. Available speeds are `r names(speed)`.

The test image may be modified to include a rectangular area of gland by R subsetting:

```{r}
img[4:8, 8:24] <- 1/speed["gland"]
```

(Reference: [Nebeker and Nelson](http://www.jultrasoundmed.org/content/31/9/1389.full) "Mean values from published sound speed reports are as follows: fat, 1478 m/s; glandular breast, 1510 m/s; benign breast tumors, 1513 m/s; and malignant breast tumors, 1548 m/s")

In R's default function for displaying images, rows index the x dimension from left to right, and columns index the y dimension from bottom to top. Thus, the "gland" occupies a rectangle at the vertical center on the left of the image. Function `gridOverlay` plots grid lines over the image.

```{r fig.align='center', fig.width=4.4}
tix <- (0:31)*attr(img,"spacing")
image(tix, tix, img, asp=ncol(img)/nrow(img), main="Test image with 'gland'\nand overlay", col=c("blue", "gray"), xlab="x (mm)", ylab="y (mm)")
gridOverlay(img)
```

A probe is represented by an array of equally spaced transducers. Its alignment is specified in the coordinate system of an associated test image. ??allow z??
